---
title: "Inference of the Tooth Growth Data Set"
author: "Travis Fell"
date: "Thursday, June 18, 2015"
output: html_document
---
#Overview
This analysis puts into action tools and techniques from the Exploratory Data Analysis class as well as hypothesis testing from the Statistical Inference class using the ToothData dataset in R. 

#Load and Explore the Data
First, we'll load the datasets package and other packages we will use in this analysis (this analysis assumes these packages have already been downloaded and instatlled).  
```{r}
library(dplyr)
library(ggplot2)
library(datasets)
data(ToothGrowth)
```
Now, we'll do some basic exploratory analysis just to get a feel for this data set. 
```{r}
dim(ToothGrowth)
ToothGrowth[!complete.cases(ToothGrowth)]
head(ToothGrowth)
View(ToothGrowth)
lenrange <- range(ToothGrowth$len)
str(ToothGrowth)
summary(ToothGrowth$len)
```
Digging a little deeper, let's look at the distribution of tooth lengths (or the purpose of this analysis, I will assume length is in microns), the number of observations by supplement delivery method and dose and the average tooth length by method and dose.
```{r}
lenave <- mean(ToothGrowth$len)
hist(ToothGrowth$len, main = "Overall Distribution of Tooth Length", xlab = "Length in Microns")
abline(v = lenave , col = "red")
nrow(subset(ToothGrowth, supp == "OJ"))
nrow(subset(ToothGrowth, supp == "VC"))
unique(ToothGrowth$dose)
nrow(subset(ToothGrowth, dose == 0.5))
nrow(subset(ToothGrowth, dose == 1.0))
nrow(subset(ToothGrowth, dose == 2.0))
```
So, the exploration of this data set has revealed the following: 
- The data set is 60 rows long with 3 columns, len, supp and dose
- The len columns contains varying tooth lengths 
- The overall range in tooth lengths is `r lenrange'
- The average tooth length is `r lenave`
- The supp column is a factor that contains two Vitamin C supplement delivery methods, OJ and VC
- Half the data set has OJ for the supp value, the other half has VC
- There are three distinct dosage levels: 0.5, 1.0, 2.0
- There are 20 records each for each dosage level listed above


#Summarize the Data
To further understand the tooth length data along by supplement method and dosage, we will find the average tooth length and standard deviation by supplement method then by dosage level. 
```{r}
growthBySupp <- summarize(group_by(ToothGrowth, supp), mean(len), sd(len))
colnames(growthBySupp) <- c("supp", "meanlen", "sdlen")
growthByDose <- summarize(group_by(ToothGrowth, dose), mean(len), sd(len))
colnames(growthByDose) <- c("supp", "meanlen", "sdlen")
```

#Assumptions

#Testing Tooth Growth
The data shows tooth growth by two dimensions: delivery method and doses. Since the sample sizes are relatively small, we'll use the T distribution to compare our hypotheses. 

The first hypothesis we'll test is whether or not the supplemental delivery method OJ (orange juice) has a greater impact on tooth growth than VC (absorbic acid). 
```{r}
(growthBySupp[1,2] - growthBySupp[2,2]) + c(-1,1) * qt(.975, 58) * (growthBySupp[1,3]^2/30 + growthBySupp[2,3]^2/30)^.5
```


#Conclusion
