---
title: "Statistical Inference Course Project"
author: "Travis Fell"
date: "Thursday, June 11, 2015"
output: html_document
---
#Overview

#Simulations
This following R code creates the exponential distribution, finds the mean and variance for the distribution, pulls the requisite number of samples, finds the distribution of all those sample means and variances for comparison to the theoretical mean and variances later. 

First, let's create the exponential distribution and plot it out. 
``` {R}
exp <- rexp(1000, .02)
hist(exp, main = "Exponential Distribution of 1000 Simulations")
```
This distribution starts high and "ski slopes" down toward zero. Now let's look at the mean and variance for this distribution.
```{r}
exp_mu <- mean(exp)
exp_var<- (1/.2)^2
```
The overall population mean, exp_mu equals `r exp_mu` and the population variance, exp_var equals `r exp_var`. 

With the parameters of the overall population established, let's take 1000 averages from 40 exponential samples then calculate the mean and variance of the distribution of those samples. 
```{r}
exp_samp <- NULL
for(i in 1:1000) exp_samp <- c(exp_samp, mean(rexp(40, .2)))
hist(exp_samp, main = "Dist of 1K Ave from 40 exponential variables")
```

The distribution of the sample is starting to look more like a normal distribution. Now, let's calculate the mean and variance for the distribution of the samples.
```{r}
exp_samp_mean <- mean(exp_samp)
exp_samp_var <- var(exp_samp)
```
The sampling mean equals `r exp_samp_mean` and the sampling variance equals `r exp_samp_var`. 

#Sample Mean Vs. Theoretical Mean

#Sample Variance Vs. Theoretical Variance

#Distribution


